---
title: 'Example Report in RMarkDown'
date: 'STA302/1001 Autumn 2017'
output: 
  pdf_document
---

<!-- You need to know R but not RMarkDown. Load this file in RStudio (and click 'Knit') only if you are interested. -->

\newcommand{\var}{\text{var}} <!-- You can define new commands just like in Latex -->


**First Heading: Simple example of a sentence**

The sum of 4 and 5 is calculated by an R command in this document to be `r 4+5`.

\vspace{1cm}  <!-- Again, just like Latex -->

**Second example**

Here is some more R code embedded in the document!

```{r}
x <- c(2.1,-1,0,4) # c = combine
y <- rt(4,6) # Four observations from a t_6 distribution
stdDev <- sd(x)
mean(y)
```

So, using more R commands, the average of the $t$-distributed sample turns out to be $\sim$ `r round(mean(y),2)`, and we have $\var(x) \approx `r round(stdDev^2,3)`$.

Data can be stored in a "data frame", which is sort of like a matrix with rows and/or columns labelled:

```{r}
z <- data.frame(x,y)
print(z)
```


\newpage  

**Third example: Visualization**

```{r}
plot(x,y,ylab="t6")
```

\vspace{1cm}  

**Fourth example: Using packages**

```{r,include=FALSE}
install.packages("dplyr", repos='http://cran.rstudio.com') # Comment this out once installed
```

```{r,eval=FALSE}
install.packages("dplyr", repos='http://cran.rstudio.com')
```

```{r,warning=FALSE,message=F}
library(dplyr) # load up the package that was just installed
filter(z,x>0) # a concise command, courtesy of dplyr
```
