\documentclass[11pt]{article}
\input{/Users/markwang/.preamble}
\begin{document}


\subsection*{NPC}

To show $X$ is NPC 

\begin{enumerate}
    \item $X\in NPC$, show that there is an efficient verifier for $X$
    \item $X$ is NP-hard, show $Y \leq_q X$ for some NP-hard problm $Y$
\end{enumerate}

\begin{example}
    \textbf{Subset-sum}($SS$) Given a finite set $S$ of positive integers and a positive integer $t$, ask if there is a subset $S' \subseteq S$ such that $\sum_{x\in S} x = t$. Prove that subset sum problem is NP-complete
    \begin{proof}
        2 steps
        \begin{enumerate}
            \item Prove \textsc{Subset-sum}$\in NPC$ Given a subset sum problem $X = (S,t)$. Let certificate $c$ be a subset $S'\subseteq S$. Let $B(X,c)$ be such that  
            \[
                B(X,c) = 
                \begin{cases}
                    yes & \text{if } \sum_{x\in c} x = t\\
                    no & \text{otherwise}
                \end{cases}
            \]
            Note $B$ can be implemented by summing up all $x\in c$ and return the sum, which will finish in polynomial time. Hence $B$ is an efficient verifier of $X$, hence  \textsc{Subset-sum}$\in NPC$
            \item Prove \textsc{Subset-sum} is NP-hard by proving $3-SAT \leq_q SS$. Start with some input to 3-SAT, and modify the input with function $f$ and make it an input $f(x)$ to SS. Proving $3-SAT\leq_q SS$ is equivalent to prove that $3-SAT$ says yes on $x$ if and only if SS says yes on $f(x)$ and $f$ is a polynomial time operation. Let $\phi$ b input to 3-SAT
            \[
                \phi  = (a_1 \lor b_1 \lor c_1) \land \cdots \land (a_r \lor b_r \lor c_r)
            \]
            where $a_i, b_i, c_i \in \{ x_1, \cdots, x_s \}$. In summary, there are $r$ clauses and $s$ variables. Let integer $\hat{x}_j = 1$ followed by $s-j$ 0s followed by $r$ digits where $k$th next digit equals 1 if $x_j$ appears in clause $c_k$, 0 otherwise. and define $\neg \hat{x}_j$ be same as previous definition but when $\neg x_j$. For $j = 1\to r$ integer $C_j = 1$ followed by $r-j$ 0s and integer $D_j = 2$ follwed by $r-j$ 0s $t = s$ 1s followed by $r$ 4s\\
            Example of $r = 3$ clauses and $s =4$ variables
            \[  
                \phi = (x_1 \lor \neg x_2 \lor \neg x_4) \land (x_2 \cup \neg x_3 \cup x_1) \land (\neg x_3 \lor x_4 \lor \neg x_2)
            \]
            Now we define $S$ 
            \begin{align*}
                 S &= \{\\
                    &\hat{x}_1 = 1000110, \neg \hat{x}_1 = 1000000, \\
                    &\hat{x}_2 = 100010, \neg\hat{x}_2 = 100101, \\ 
                    &\hat{x}_3 = 10000, \neg \hat{x}_3 = 10011, \\
                    &\hat{x}_4 = 1001, \neg\hat{x}_4 = 1100, \\ 
                    &C_1 = 100, C_2 = 10, C_3 = 1 \\
                    &D_1 = 200, D_2 = 20, D_3 = 2 \\
                    &t = 1111444 \}\\ 
            \end{align*}
            Th operation takes polynomial time. now we show $\phi$ is satisfiable if and only if $S$ has a subset whose sum is $t$
            \begin{enumerate}
                \item $=>$ Assume $\phi$ satisfiable, implying each clause must have one true literal (variable). Let
                \[
                    S' = \{ \text{numbers that correspond to the literals as constructed} \}
                \] 
                By construction, choose $S'$ such that $\sum_{x\in S'} x = 1111d_1 d_2 d_3$ where $1 \leq d_i \leq 3$ (Since at least 1 true varaible, and and at most 3 variables in a clause) Now we can select $S''\subseteq \{ C_i, D_i\}$, such that $\sum_{x\in S' \cup S''} x = t$
                \item $<=$ There is some subset $S' \subseteq S$ such that $\sum_{x\in S'} x=  t$ The truth assignment, $T(x_i)$ is true if $\hat{x}_i \in S'$ and $T(x_i)$ is false if $\neg \hat{x}_i \in S'$. The sum of all $C_i$s and $D_i$s is 333. So if $\sum_{x\in S'}x = t = 1111444$ which means we have some 1 in each of the 3 least significant bits. Then there is a truth literal in that clause, so every clause is true, and $\phi$ is true
            \end{enumerate}
        \end{enumerate}
    \end{proof}
\end{example}


\begin{theorem*}
    If $D$ is NPC, then $D\in P \iff P=NP$
    \begin{proof}
        prove if and only if
        \begin{enumerate}
            \item $=>$ $D\in P$ let $D' \in NP$ Show $D' \in P$ D is NPC, implies $D' \leq_q D$, implies $D' \in P$
            \item $<=$ $D$ is NPC implies $D\in NP = P$ implies $D\in P$
        \end{enumerate}
    \end{proof}
\end{theorem*}

\subsection*{Categories of NPC problems}


\begin{enumerate}
    \item \textbf{Packing problem} 
    \begin{enumerate}
        \item independent set
        \item set packing 
    \end{enumerate}
    \item \textbf{Covering problem}
    \begin{enumerate}
        \item vertex cover 
        \item set cover
    \end{enumerate}
    \item \textbf{Partition problem}
    \begin{enumerate}
        \item 3-dimensinoal matching
        \item graph coloring
    \end{enumerate}
    \item \textbf{Sequencing problem}
    \begin{enumerate}
        \item Hamiltonian cycle
        \item Hamiltonian paht
        \item traveling salesman
    \end{enumerate}
    \item \textbf{Numerical problems}
    \begin{enumerate}
        \item subset sum 
    \end{enumerate}
    \item \textbf{Constraint satisfaction problems}
    \begin{enumerate}
        \item 3-SAT
        \item CNF-SAT 
        \item SAT 
        \item Circuit-SAT
    \end{enumerate}
\end{enumerate}


\subsection*{Self-Reducible Problems}

Search problem, find a solution \\ 
optimization problem, find the best prossible solution 


\begin{defn*}
    \textbf{Search problem}
    A search problem is self-reducible if any efficient solution to the corresponding decision problem can be used to solve the search problem efficiently.  Search problems are typically harder than the corresponding decision problem. 
\end{defn*}

\begin{defn*}
    \textbf{Independent-Set-Decision}$(G, k)$ 
\end{defn*}

\begin{example}
    \textbf{Hamiltonian cycle} \\

    \begin{enumerate}
        \item problem is self-reducible Let $HCS$ be the search problem and $HCD$ be the decision problem. 
        \item \textbf{Correctness} $HCD(G')$ remains true for every state. So at the end, $E$ contains every edge in hamiltonian cycle of $G$. Moreover, we remove every other edge so we are left with only a hamiltonian cycle at the end. 
        \item \textbf{Complexity} Assume the time complexity of decision problem is $t(V, E)$. Then the complexity of the search problem is $O(t(V,E) + E(1 + t(V,E))) = O((E+1)t(V,E))$
    \end{enumerate}

    \begin{algorithm}[H]
        \SetKwFunction{HCS}{Hamiltonian-cycle-search}
        \SetKwFunction{HCD}{Haniltonian-cycle-decision}
        \SetKwFunction{NIL}{NIL}

        \Fn{$\HCS(G)$}{

        \If{! $\HCD(G)$}
            \Return{\NIL}\\
        }
        \For{$e\in G.E$}{
            $E' \leftarrow E \setminus \{ e\}$ \\
            $G' \leftarrow (V, E')$\\
            \If{$\HCD(G')$}{
                $E \leftarrow E'$ \\ 
            }
        }
        \Return{$E$}\\
    \end{algorithm}
\end{example}


\begin{example}
    \textbf{Vertex-Cover}

    \begin{enumerate}
        \item \textbf{Complexity} $O((V+1)t(V,E) + V\times(V+E))$
    \end{enumerate}

    \begin{algorithm}[H]
        \SetKwFunction{VCS}{Vertex-Cover-Search}
        \SetKwFunction{VCD}{Vertex-Cover-Decision}

        \Fn{$\VCS(G,k)$}{
            \If{! $\VCD(G,k)$}{
                \Return{\NIL}
            }
            $C \leftarrow \{ \}$ \\
            \For{$v \in V$}{
                \If{$k \leq 0$}{
                    break \\ 
                }
                $V' \leftarrow V \setminus \{ v\}$ \\
                $E' \leftarrow E \setminus \{ (v,w) | (v,w)\in E\}$ \\ 
                $G' \leftarrow (V', G')$\\
                \If{$\VCD(G', k-1)$}{
                    $v\leftarrow v'$\\ 
                    $E \leftarrow E'$\\
                    $C \leftarrow C \cup \{ v\}$\\
                    $k \leftarrow k-1$ \\ 
                }
            }
        }
    \end{algorithm}
\end{example}




\subsection*{Optimization Problem} 

Independent set, given $G$ find an independent set of maximum size. use binary search to find left/right partition with independent-set-decision


\subsection*{handling intractable problems}
It is not possible to solve NP-hard problems exactly and in polynomial time (unless $P = NP$)

\begin{defn*}
    \textbf{Approaches}
    \begin{enumerate}
        \item \textbf{Sacrifice efficiency} NP-hard problems, generic problems that accounts for worst case problems. In reality, the inputs may be much better. For example find independent set on a tree
        \item \textbf{Sacrifice exactness} For optimization problems, it is okay to find a good enough solution compared to the best possible solution. We can use approximation algorithms.
    \end{enumerate}
\end{defn*}

\begin{example}
    \textbf{Vertex-cover} 
    \begin{enumerate}
        \item Find a set of vertices of minimum possible sizse of every edge in $G$ passes through one of these vertices.
        \item claim $|S| \leq 2 \times OPT$
        \begin{proof}
            Any vertex cover of $G$ must include at least one of the two vertices connecting an edge in $S$. Also, in $S$, all edges are disjoint, i.e. with no end points in common. so $OPT \geq |S| / 2$
        \end{proof}
        \item In general, how can we comput approximation ratio if we dont know what the optimal solution is? use a lower bound instead. For other value LB that is easy to compute, show that $LB \leq OPT$ and solution $A\leq ratio \times LB$
    \end{enumerate}
    \begin{algorithm}[H]
        \Fn{}{
            $S \leftarrow \{ \}$ \\
            \For{$e = (u,v) \in E$}{
                $S \leftarrow S \cup \{v,u \}$\\
                $E'\leftarrow \{ (u,w) | (u,w)\in E \} \cup \{ (v,w) | (v,w)\in E\}$\\
                $E \leftarrow E'$ \\
            }
            \Return{$S$}\\ 
        }
    \end{algorithm}
\end{example}





\end{document}
